\boldmath Supersymmetric Dirac-Born-Infeld Axionic Inflation and Non-Gaussianity

### 2018-09-23 SUSY_DBI_referee_reply.tex

\noindent {\it Reply\\ 
We agree that our references to the analysis of non-Gaussianity from the Planck experiment needs updating. We have
done so in the revised version.  The changes made appear after Eq.(5.12) and at the end of section 6. We have included
the reference to 1502.01592 and specified the current limits on $f_{NL}^{equil}$.
}

4) I don't understand eq.(6.1). The conventional way to find the end
of inflation is to find the moment at which $\ddot a = 0$ (second derivative of the
scale factor with respect to time). The number of e-folds during inflation is
typically defined from $a(N ) \equiv a_{end} e^{-N}$ , which gives  $\frac{dN}{dt} = -H$ (where $a_{end}$ 
is the value of the scale factor at the end of inflation, and H is the Hubble
rate). In this way, $N = 0$ at the end of inflation, while N is positive at
early times during inflation. How is eq. (6.1) used to determine the end of
inflation ?  What is the purpose of setting $\theta_n = \frac{1}{16}$?  Why don't the authors 
find from their numerical evolutions the point at which $\ddot a = 0$ ?\\
\noindent {\it Reply
}

5) The scan shown in Figure 1 contains no real information, unless one
specifies what is varied among the different points. The only information I
could find on this regard is we allow $\alpha_1$ to vary at the end of page 12.
How is it varied ? (in which range ? Is it a uniform sampling in linear units
? In log units ?) What is chosen for the other parameters of the model
? (?,?,A,T). How is the correct normalization of the scalar perturbations
enforced? (this is typically done by fixing the overall scale of the scalar
potential). Are subleading terms in $\frac{1}{T}$  important ? The scan strategy should T
be completely described, so that any reader, if interested, could precisely reproduce those figures.\\

\noindent {\it Reply\\
We agree that the distribution used to generate the dataset from Figs.(1, 2) is important to be
mentioned in the paper. We added a complete specification to the caption of Fig.(1).
}

@@

6) The paper motivates the model from the trans-Planckian problem of axion inflation (the fact that we must take the axion scale 
$f > M_p$ in the original potential of natural inflation). It would be very interesting to provide also a scan of the values of f obtained for the various realizations presented in figures 1 and 2. Does this scan extend in the sub-Planckian region ? (the scan should be easily done, once the above questions 2.4 and 5 are answered).\\

\noindent {\it Reply\\
We agree with the suggestion of the referee. We have added a new figure 3 where we have presented 
 a scatter plots of $f$ vs $n_s$ and $f$ vs $r$ for the data sets of figures 1 and 2.
 One can see that $20\%$ of experimentally consistent points satisfy $f < M_\text{P}$.

### 2018-09-23 SUSY_DBI.tex
Plot of Decay Constants for Referee #6 (#13)
* Closes #10.
* Added fig4 with plots of $f$ vs $n_s$ and $f$ vs $r$.
* Added a note to the main text about fig4.
* Updated point 6 in the referee reply.

  f&= f_0 - \frac{1}{\lambda} \sum_{k} \frac{kA_k}{2^{k/2-1} } f_0^{k-3} + \cdots  \nonumber\\

@@

@@

Numerical simulations performed consist of two parts: a large $0.5 \times 10^6$ points simulation in which $\alpha_2 = \alpha_3 = 0$, and a small simulation with $\alpha_2$ and $\alpha_3$ perturbed around an experimentally consistent point with the largest non-Gaussianity from the first simulation.
  The evolution of the inflaton field $a_-$ in the large simulation of Figs.~(\ref{fig1}) and (\ref{fig2}) is obtained by taking the effective axion Lagrangian Eqs.~(\ref{eq.axionLagrangian.I}) and (\ref{eq.axionLagrangian.II}), deriving Euler-Lagrange equation and Friedmann Eq.~(\ref{fried2}) using computer algebra, and solving them numerically. The integration is stopped when the ratio of the time derivative ($\dot n\left(t\right)$) of the e-foldings count to its time-average ($n\left(t\right) / t$) passes a threshold. Specifically,
  \begin{equation} 
  	\frac{\dot n \left(t\right)}{n\left(t\right) / t} < \theta_n,
  \end{equation} 
  where the threshold $\theta_n$ is chosen to be much less than one, and specifically
   in the analysis here it is chosen as $\theta_n = 1 / 16$. Increasing this parameter might stop integration prematurely, whereas decreasing it will worsen performance.
  An analytic expression similar to Eq.~(\ref{eq.axionLagrangian.II}), however,  cannot be derived  for  non-zero $\alpha_2$ and $\alpha_3$. In this case,  equations for the auxiliary fields $F_k$ are  solved numerically,  and the Lagrangian in terms of $\phi$ (Eq.~(\ref{Lag.23})) is
  also evaluated numerically for each set of values for ($a_-$, $\dot a_-$).  
   This process is significantly slower, therefore, in this case the simulations are only done for $29$ sets of parameter values as can be seen in Fig.~(\ref{fig3}). 
    The time of horizon exit is then found by counting $N_\text{pivot}$ e-foldings back from the end of integration, where $N_\text{pivot}$ is varied between $50$ and $60$. Note, the time of horizon exit depends on $\theta_n$, however, we find that due to $\dot n$ being small near the end of integration, the dependence is weak, and it affects the results less than the experimental uncertainty of $N_\text{pivot} \in \left[50, 60\right]$.{}

  Finally, we evaluate the experimental observables at the time of horizon exit using equations from section \ref{sec5} and reusing Euler-Lagrange and Friedmann equations to compute higher time derivatives of the Lagrangian, pressure and density. The observables computed are the ratio $r$ of the power spectrum of tensor to scalar perturbations, the spectral indices of scalar $n_s$ and tensor $n_t$ perturbations, 
 speed of sound $c_s$ 
  and the non-Gaussianity amplitude $f_\text{NL}$.
    In the $0.5 \times 10^6$ points Monte Carlo analysis we allow $\alpha_1$ to vary
  (full distribution is specified in Fig.~(\ref{fig1}))
  and search for solutions that satisfy the experimental constraints. Thus the
  current experimental limits from Planck experiment at $k_0=0.05\, {\rm Mpc}^{-1}$ are as follows~\cite{Adam:2015rua,Ade:2015lrj,Array:2015xqh}
  \begin{align}
  	n_s& = 0.9645 \pm 0.0049\, (68\% {\rm CL})\,, \non
  	r & <0.07\, (95\% {\rm CL})\,.
  	\label{data}
  \end{align}
  and the current experimental constraint on $c_s$ is~\cite{Ade:2015lrj}
  \begin{align}
  	c_s\geq 0.087~~~({\rm at ~}95\% ~CL)\,.
  	\label{sound}
  \end{align}
  The results of the analysis for the two field supersymmetric DBI discussed above
  are presented in Figs. (\ref{fig1}) and (\ref{fig2}). The left panel of Fig. (\ref{fig1}) gives
  a plot of $r$ vs $n_s$. Here the region enclosed by the blue line is the one allowed by experiment in the $2\sigma$ range.
  The Monte Carlo analysis shows that the experimentally allowed region is well populated by the parameter points of the model.
  One interesting feature of the analysis is that compared to the
  supersymmetric axion models discussed in~\cite{Nath:2017ihp}
  where $r$ was extremely small, here $r$ has significantly
  larger values. Thus the largeness of $r$ discriminates this class of supersymmetric DBI models from the models
  of \cite{Nath:2017ihp}.
  The right panel of Fig. (\ref{fig1}) gives a plot of $f^{equil}_{NL}$ vs $\alpha_1$. We see a significant sensitivity
  of $f^{equil}_{NL}$ to $\alpha_1$. However, overall $f^{equil}_{NL}$ is typically small and mostly lies below 0.5. The effect of $\alpha_2$ and $\alpha_3$ for the point from Fig.~\ref{fig1} with largest non-Gaussianity is shown on Fig.~\ref{fig3}. One can see that although non-Gaussianity $f_\text{NL}$ is sensitive to $\alpha_2$ and $\alpha_3$, it still lies below 0.5. This level
  of non-Gaussianity appears too low for observation in the near future.

@@

>>>>

    \begin{align} \label{eq.f.series}
  f&= f_0 - \frac{2}{3\lambda} \sum_{k} \frac{kA_k}{2^{k/2-1} } f_0^{k-3} + \cdots  \nonumber\\

@@

  \begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{figs/fig4.pdf}
    \caption{Left panel: A plot of axion decay constant $f$ vs scalar spectral index $n_s$ for the
      same data set as in Fig.~(\ref{fig1}).
          Right panel: The same as the left panel except $f$ is plotted vs tensor to scalar ratio
      $r$.
          Note that $20\%$ of experimentally consistent (green) points in the figure satisfy
      $f<M_\text{P}$. Also note that we vary $f$ explicitly in this paper, which implicitly sets
      the value of $\lambda$, as can be seen from Eq.~(\ref{eq.f.series}).}
    \label{fig4}
  \end{figure}

@@

Numerical simulations performed consist of two parts: a large $0.5 \times 10^6$ points simulation in which $\alpha_2 = \alpha_3 = 0$, and a small simulation with $\alpha_2$ and $\alpha_3$ perturbed around an experimentally consistent point with the largest non-Gaussianity from the first simulation.
  The evolution of the inflaton field $a_-$ in the large simulation of Figs.~(\ref{fig1}), (\ref{fig2}), and (\ref{fig4}) is obtained by taking the effective axion Lagrangian Eqs.~(\ref{eq.axionLagrangian.I}) and (\ref{eq.axionLagrangian.II}), deriving Euler-Lagrange equation and Friedmann Eq.~(\ref{fried2}) using computer algebra, and solving them numerically. The integration is stopped when the ratio of the time derivative ($\dot n\left(t\right)$) of the e-foldings count to its time-average ($n\left(t\right) / t$) passes a threshold. Specifically,
  \begin{equation} 
  	\frac{\dot n \left(t\right)}{n\left(t\right) / t} < \theta_n,
  \end{equation} 
  where the threshold $\theta_n$ is chosen to be much less than one, and specifically
   in the analysis here it is chosen as $\theta_n = 1 / 16$. Increasing this parameter might stop integration prematurely, whereas decreasing it will worsen performance.
  An analytic expression similar to Eq.~(\ref{eq.axionLagrangian.II}), however,  cannot be derived  for  non-zero $\alpha_2$ and $\alpha_3$. In this case,  equations for the auxiliary fields $F_k$ are  solved numerically,  and the Lagrangian in terms of $\phi$ (Eq.~(\ref{Lag.23})) is
  also evaluated numerically for each set of values for ($a_-$, $\dot a_-$).  
   This process is significantly slower, therefore, in this case the simulations are only done for $29$ sets of parameter values as can be seen in Fig.~(\ref{fig3}). 
    The time of horizon exit is then found by counting $N_\text{pivot}$ e-foldings back from the end of integration, where $N_\text{pivot}$ is varied between $50$ and $60$. Note, the time of horizon exit depends on $\theta_n$, however, we find that due to $\dot n$ being small near the end of integration, the dependence is weak, and it affects the results less than the experimental uncertainty of $N_\text{pivot} \in \left[50, 60\right]$.{}

  Finally, we evaluate the experimental observables at the time of horizon exit using equations from section \ref{sec5} and reusing Euler-Lagrange and Friedmann equations to compute higher time derivatives of the Lagrangian, pressure and density. The observables computed are the ratio $r$ of the power spectrum of tensor to scalar perturbations, the spectral indices of scalar $n_s$ and tensor $n_t$ perturbations, 
 speed of sound $c_s$ 
  and the non-Gaussianity amplitude $f_\text{NL}$.
    In the $0.5 \times 10^6$ points Monte Carlo analysis we allow $\alpha_1$ to vary
  (full distribution is specified in Fig.~(\ref{fig1}))
  and search for solutions that satisfy the experimental constraints. Thus the
  current experimental limits from Planck experiment at $k_0=0.05\, {\rm Mpc}^{-1}$ are as follows~\cite{Adam:2015rua,Ade:2015lrj,Array:2015xqh}
  \begin{align}
  	n_s& = 0.9645 \pm 0.0049\, (68\% {\rm CL})\,, \non
  	r & <0.07\, (95\% {\rm CL})\,.
  	\label{data}
  \end{align}
  and the current experimental constraint on $c_s$ is~\cite{Ade:2015lrj}
  \begin{align}
  	c_s\geq 0.087~~~({\rm at ~}95\% ~CL)\,.
  	\label{sound}
  \end{align}
  The results of the analysis for the two field supersymmetric DBI discussed above
  are presented in Figs. (\ref{fig1}), (\ref{fig2}), and (\ref{fig4}). The left panel of Fig. (\ref{fig1}) gives
  a plot of $r$ vs $n_s$. Here the region enclosed by the blue line is the one allowed by experiment in the $2\sigma$ range.
  The Monte Carlo analysis shows that the experimentally allowed region is well populated by the parameter points of the model.
  One interesting feature of the analysis is that compared to the
  supersymmetric axion models discussed in~\cite{Nath:2017ihp}
  where $r$ was extremely small, here $r$ has significantly
  larger values. Thus the largeness of $r$ discriminates this class of supersymmetric DBI models from the models
  of \cite{Nath:2017ihp}.
  The right panel of Fig. (\ref{fig1}) gives a plot of $f^{equil}_{NL}$ vs $\alpha_1$. We see a significant sensitivity
  of $f^{equil}_{NL}$ to $\alpha_1$. However, overall $f^{equil}_{NL}$ is typically small and mostly lies below 0.5. The effect of $\alpha_2$ and $\alpha_3$ for the point from Fig.~\ref{fig1} with largest non-Gaussianity is shown on Fig.~\ref{fig3}. One can see that although non-Gaussianity $f_\text{NL}$ is sensitive to $\alpha_2$ and $\alpha_3$, it still lies below 0.5. This level
  of non-Gaussianity appears too low for observation in the near future.

@@

  Finally, Fig.~(\ref{fig4}) demonstrates that it is indeed possible to produce experimentally
  consistent inflation with $f < M_\text{P}$ in our model.
